---
title: "Covid91 vaccine study Final2021"
author: "Emann Malik"
date: "May 7, 2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

```{r include = FALSE}
FinalData= select(FinalData, ! is.numeric)
FinalData=purrr::map(FinalData, as.factor)
FinalData = data.frame(FinalData)
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.
  
# Males
  
```{r}
males <- subset(FinalData, sex == "guy")
``` 
  
## Introduction
  
$H_0$ The covid91 vaccine is effective in males.  
$H_A$ The covid91 vaccine is not effective in males.  
  
## Descriptive Results
  
### Graphical Descriptive Results
  
```{r}
barchartGC(~infected + treatment, data=males)
barchartGC(~infected + treatment, data=males, type="percent")
```
  
The bar chart above shows that, out of all the males in this trial who took the covid91 vaccine, almost 60% got infected with covid and, out of all the males in this trial who took the placebo, over 40% did not get infected with covid. That means the amount of males who took the vaccine and got sick is 1 and a half times as much as those who took the placebo and got sick! We are already starting to see evidence that the covid91 vaccine may not be effective for males. 
  
### Numerical Descriptive Results
  
```{r}
table1 <- xtabs(~treatment + infected, data=males)
table2 <- xtabs(~infected + treatment,data=males)
rowPerc(table1)
colPerc(table1)
```
  
The first table above shows that more males who took the covid91 vaccine got infected with covid (3.72) than the males who took the placebo and got infected (2.55). Likewise, the second table above also displays similar proportions; more of the infections came from the vaccinated males (58.16) than from the males who took the placebo (41.84). Once again, it does not look like the covid91 vaccine is a good hope for males. 
  
## Inferential Results
  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
fisher.test(table2)
```
  
$H_0$ The covid91 vaccine is effective in males.  
$H_A$ The covid91 vaccine is not effective in males.  
  
The Chi-squared test above displays a p-value of 0.0006181 while Fisher's Exact Test displays a p-value of 0.0005801. Both of these values indicate some sort of significance. They, both representing the chance that we will get the same results just received if the null hypothesis is true, are very small values.
  
So, based on these values, we can reject the null hypothesis in favor of the alternative: the covid91 vaccine is not effective in males. 
  
## Conclusion
  
Because our p-values for the male subgroup are too small to show a significant chance that we will get the same results just received if the null hypothesis is true, we reject it in favor of the alternative: the covid91 vaccine is not effective in males.
  
The covid91 vaccine is clearly not effective in this subgroup. I would not recommend males to take this vaccine.
  
# Females
  
```{r}
females <- subset(FinalData, sex == "gal")
``` 
  
## Introduction
  
$H_0$ The covid91 vaccine is not effective in females.  
$H_A$ The covid91 vaccine is effective in females.  
  
## Descriptive results
  
### Graphical Descriptive Results
  
```{r}
barchartGC(~infected + treatment, data=females)
barchartGC(~infected + treatment, data=females, type="percent")
```
  
The bar chart above is already showing us proof that the covid91 vaccine is promising for the female subgroup; more than 60% of the females in the study who became infected with the virus took the placebo while almost 40% of the females in the study who became infected with the virus took the vaccine. That means more sick females came from the placebo group than the vaccinated group.
  
### Numerical Descriptive Results
  
```{r}
table1 <- xtabs(~treatment + infected, data=females)
table2 <- xtabs(~infected + treatment,data=females)
rowPerc(table1)
colPerc(table1)
```
  
The first table above shows us that, out of all the females involved in the study who got infected with covid, a larger proportion of them took them had taken the placebo (8.67) than the vaccine (4.44). The second table above shows us similar information; less vaccinated females got sick than the female placebo group. Once again, it appears that the covid91 vaccine is effective in females.
  
## Inferential Results
  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
fisher.test(table2)
```  
  
$H_0$ The covid91 vaccine is not effective in females.  
$H_A$ The covid91 vaccine is effective in females.  
  
The Chi-squared test above displays a p-value of 0.0000000000000003626 while Fisher's Exact Test displays a p-value that is less than 0.00000000000000022. Both of these values definitely indicate some sort of significance. They, both representing the chance that we will get the same results just received if the null hypothesis is true, are astoundingly small values.  
  
These p-values stack up high against the null hypothesis. So, based on these values, we can reject the null hypothesis in favor of the alternative: the covid91 vaccine is effective in females.
  
## Conclusion
  
Because our p-values for the female subgroup are much too small to show a significant chance that we will get the same results just received if the null hypothesis is true, we reject it in favor of the alternative: the covid91 vaccine is effective in females.
  
The covid91 vaccine is clearly effective in this subgroup. I would definitely recommend females to take this vaccine.  
  
# Drug Users
  
```{r}
drugUsers <- subset(FinalData, DrugUser == "yes")
```   
  
## Introduction
  
$H_0$ The covid91 vaccine is not effective in drug users.  
$H_A$ The covid91 vaccine is effective in drug users.  
  
## Descriptive results
  
### Graphical Descriptive Results
  
```{r}
barchartGC(~infected + treatment, data=drugUsers)
barchartGC(~infected + treatment, data=drugUsers, type="percent")
``` 
  
The bar chart above shows that, out of all the drug users involved in this study who got infected with covid, almost 70% had taken the placebo and nearly 40% had taken the covid91 vaccine. That means there were 1.75 times as many sick placebo drug users as there were sick vaccinated drug users. If we look at the frist group of two bars in the plot, the right one representing the amount of placebo infections appears to be just a little over twice the size of the left bar representing the vaccine infections! It is beginning to appear that the covid91 vaccine is working well in drug users.
  
### Numerical Descriptive Results
  
```{r}
table1 <- xtabs(~treatment + infected, data=drugUsers)
table2 <- xtabs(~infected + treatment,data=drugUsers)
rowPerc(table1)
colPerc(table1)
```  
  
The first table above shows us that the proportion of drug users who took the placebo and got sick (6.46) is twice as much as the proporation of drug users who took the vaccine and got sick (6.46). Likewise, the second table underneath it also shows similar data; the proportion of infected placebo drug users (66.67) outnumbers the proportion of infected drug users who were vaccinated (33.33) by twice as much. Once again, it appears that the covid91 vaccine is doing its job for the drug users. 
  
## Inferential Results
  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
fisher.test(table2)
``` 
  
$H_0$ The covid91 vaccine is not effective in drug users.  
$H_A$ The covid91 vaccine is effective in drug users.  
  
The Chi-squared test above displays a p-value of  0.003028 while Fisher's Exact Test displays a p-value of 0.002273. Both of these values indicate some sort of significance. They, both representing the chance that we will get the same results just received if the null hypothesis is true, are small values.
  
So, based on these values, we can reject the null hypothesis in favor of the alternative: the covid91 vaccine is effective in drug users.
  
## Conclusion
  
Because our p-values for the drug user subgroup are too small to show a significant chance that we will get the same results just received if the null hypothesis is true, we reject it in favor of the alternative: the covid91 vaccine is effective in drug users.
  
The covid91 vaccine is clearly effective in this subgroup. I would recommend drug users to take this vaccine.  
  
# The LGBTQ Community  
  
```{r}
LGBTQ <- subset(FinalData, LGBTQ == "gay")
``` 
  
## Introduction
  
$H_0$ The covid91 vaccine is effective in the LGBTQ community.  
$H_A$ The covid91 vaccine is not effective in the LGBTQ community.  
  
## Descriptive results
  
### Graphical Descriptive Results
  
```{r}
barchartGC(~infected + treatment, data=LGBTQ)
barchartGC(~infected + treatment, data=LGBTQ, type="percent")
```  
  
The bar chart above shows that, out of all of the people involved in the study who identify as LGBTQ and became infected with covid, almost 70% had taken the covid91 vaccine and about 30% had taken the placebo. That means the group who got sick after taking the vaccine is much more than twice the size of the group who got sick after taking the placebo! We are starting to see some evidence that this vaccine may not be the best choice for the LGBTQ community.   
  
### Numerical Descriptive Results
  
```{r}
table1 <- xtabs(~treatment + infected, data=LGBTQ)
table2 <- xtabs(~infected + treatment,data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```  
  
The first table above shows that the proportion of LGBTQ study participants who received the covid91 vaccine and got infected with the virus - 7.97 - is more than twice the amount of the LQBTQ study participants who received the placebo and got infected with the virus - 3.36. The second table shows similar information; 70.69% of vaccinated LGBTQ participants came down with the virus as opposed to 29.31% of LGBTQ participants who took the placebo. Once more, we can see that the covid91 vaccine is not that effective in the LGBTQ subgroup.
  
## Inferential Results
  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
fisher.test(table2)
```
  
$H_0$ The covid91 vaccine is effective in the LGBTQ community.  
$H_A$ The covid91 vaccine is not effective in the LGBTQ community.  
  
The Chi-squared test above displays a p-value of 0.0000105 while Fisher's Exact Test displays a p-value of 0.000005741. Both of these values indicate some sort of significance. They, both representing the chance that we will get the same results just received if the null hypothesis is true, are very small values.
  
So, based on these values, we can reject the null hypothesis in favor of the alternative: the covid91 vaccine is not effective in the LGBTQ community.
  
## Conclusion
  
Because our p-values for the male subgroup are too small to show a significant chance that we will get the same results just received if the null hypothesis is true, we reject it in favor of the alternative: the covid91 vaccine is not effective in the LGTBQ community.
  
The covid91 vaccine is clearly not effective in this subgroup. I would not recommend people who identify as LGBTQ to take this vaccine.
  
# Overall Report Conclusion  
  
In this report, we used the dplyr function subset to take out the male, female, drug user, and LGBTQ subgroups from our FinalData data set. Then, we considered each subgroup separately, performing a graphical, numerical, and inferential analysis for each one. 
  
To conclude, we have found that the covid91 vaccine is not effective in males, is effective in females, is effective in drug users, and is not effective in the LGBTQ community. 
  
I would only recommend females and drug users to take the covid91 vaccine. I would not recommend males or the LGBTQ community to take it.
  